---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# remarker

<!-- badges: start -->
<!-- badges: end -->

**Remarker** is an R package for reading in Markdown files, applying transformations, and emitting R code to 

With remarker, you can write a Shiny application using Markdown, without needing the complexity of rmarkdown's `runtime: shiny` or `shiny_prerendered`. The transformations will allow the application author to modify and restructure objects in the document before they are transformed to HTML. 


## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("rstudio/remarker")
```

## Overview

Remarker operates in three phases:

* `md_ast()`: Reads in the Markdown document and converting to the Pandoc AST. This requires Pandoc to be installed.
* `ast_ir()`: Converts the Pandoc AST to remarker's intermediate representation (IR), which contains the same information, but in a format that's more concise and easier to manipulate.
* `ir_transform()`: Applies user-defined transformations on the IR.
* `ir_r()`: Converts the IR to R code which uses htmltools to do the HTML markup. This R code can then be executed.

## Examples

### A basic Shiny application

A Markdown file with a basic Shiny application is in [inst/examples/shiny_basic.md](inst/examples/shiny_basic.md). Here's how to read it in and convert it to the R code:

```{r}
library(remarker)
md_file <- system.file("examples/shiny_basic.md", package = "remarker")

x <- md_ast(md_file) # Convert Markdown to Pandoc AST
x1 <- ast_ir(x)      # Convert to remarker IR
x2 <- ir_r(x1)       # Convert IR to R code
```

At this point, we can look at the resulting R code by simply printing `x2`:

```{r}
x2
```

And this can be run as a Shiny application with `app_r()`:

```{r eval=FALSE}
app_r(x2)
```


### A Shiny application with transformations

A Shiny application with transformations is in [inst/examples/shiny_transform.md](inst/examples/shiny_transform.md). This has two notable parts.

The first is that the layout is defined as follows:

````
``` {.layout .grid}
|      |250px   |1fr        |1fr        |
|:-----|:-------|:----------|:----------|
|100px |header  |header     |header     |
|1fr   |sidebar |main_panel |main_panel |
|1fr   |sidebar |main_panel |main_panel |
```
````

This defines a grid layout. A transformer function converts this into the appropriate HTML and CSS for the layout.

The second part of interest is the main panel. It has the class `.card`. Remarker has a transformer registered for sections with this class, which is explained in the section itself:

````
## Main panel {#main_panel .card}

This section is transformed into a **card**. The section heading is displayed with `H3` (instead of the normal `H2`), and the entire content is wrapped in a `wellPanel()`.

``` {.r .ui}
plotOutput("plot")
```

``` {.r .server}
output$plot <- renderPlot({
  plot(seq_len(input$x))
})
```
````

The R code for the transformed section looks like this:

``` r
  mainPanel(
    tagList(wellPanel(
      h3(id = "main_panel", "Main panel"),
      p(
        "This section is transformed into a ",
        tags$b("card"),
        ". The section heading is displayed with ",
        code("H3"),
        " (instead of the normal ",
        code("H2"),
        "), and the entire content is wrapped in a ",
        code("wellPanel()"),
        "."
      ),
      {plotOutput("plot")},
      NULL
    ))
  )
```

To test it out:

```{r eval=FALSE}
md_file <- system.file("examples/shiny_transform.md", package = "remarker")

x <- md_ast(md_file)   # Convert Markdown to Pandoc AST
x1 <- ast_ir(x)        # Convert to remarker IR
x2 <- ir_transform(x1) # Apply transformations
x3 <- ir_r(x2)         # Convert IR to R code

app_r(x3)
```


### Static HTML output

It's even possible to convert the IR to static HTML, without Shiny.

```{r eval=FALSE}
md_file <- system.file("examples/simple_md.md", package = "remarker")
x <- md_ast(md_file)
x1 <- ast_ir(x)
x2 <- ir_r(x1)

# Display output in the Viewer panel
x2 %>% parse(text = .) %>% eval() %>% htmltools::browsable()
```
