---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# remarker

<!-- badges: start -->
<!-- badges: end -->

**Remarker** is an R package for reading in Markdown files, applying transformations, and emitting R code to 

With remarker, you can write a Shiny application using Markdown, without needing the complexity of rmarkdown's `runtime: shiny` or `shiny_prerendered`. The transformations will allow the application author to modify and restructure objects in the document before they are transformed to HTML. 


## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("wch/remarker")
```

## Overview

Remarker operates in three phases:

* `md_ast()`: Reads in the Markdown document and converting to the Pandoc AST. This requires Pandoc to be installed.
* `ast_ir()`: Converts the Pandoc AST to remarker's intermediate representation (IR).
* `ir_transform()`: Applies user-defined transformations on the IR.
* `ir_r()`: Converts the IR to R code which uses htmltools to do the HTML markup. This R code can then be executed.

## Examples

### A basic Shiny application

A Markdown file with a basic Shiny application is in [inst/examples/shiny_basic.md](inst/examples/shiny_basic.md). Here's how to read it in and convert it to the R code:

```{r}
library(remarker)
md_file <- system.file("examples/shiny_basic.md", package = "remarker")

x <- md_ast(md_file) # Convert Markdown to Pandoc AST
x1 <- ast_ir(x)      # Convert to remarker IR
x2 <- ir_r(x1)       # Convert IR to R code
```

At this point, we can look at the resulting R code by simply printing `x2`:

```{r}
x2
```

And this can be run as a Shiny application with `run_r()`:

```{r eval=FALSE}
run_r(x2)
```



It's even possible to convert the IR 

```{r eval=FALSE}
z <- md_ast("temp/simple_md.Rmd")
z1 <- ast_ir(z)
z2 <- ir_r(z1)
z2 %>% parse(text = .) %>% eval() %>% htmltools::browsable()
```
